---
title: "Data Science"
subtitle: "Ryan Clark"
format: html
editor: visual
---

## What do you think being a data scientist is about?

Being a data scientist seems to mean being a jack of all trades with respect to understanding data and making data-driven business decisions. They have a hand in computer science/programming, engineering, machine learning, and probability/statistics. This translates to several responsibilities, including data wrangling, computing, interpreting, visualizing, and communicating findings from data. Data scientist often work in industry, as opposed to academia, so the context of their data is often regarding the business and market they work within (e.g. finance, technology, healthcare, etc).

## What differences/similarities do you see between data scientists and statisticians?

I believe this is a topic of ongoing debate, particularly as you start getting into the finer details of definitions. It's not unusual for statisticians nowadays to be employed with the title of "data scientist". This also gets more complicated when you consider the knowledge and duties of master's vs doctorate-level statisticians. That said, we might generalize the following:

1. **Similarities:** Both professions make use of math and statistics, especially applied statistics. They are comfortable with operations involving data (e.g. cleaning, managing, manipulating, etc). And they use many of the same software and programming languages (e.g. R, Python). 

2. **Differences:** The training and expertise is perhaps the greatest difference. Given the recency of data science as a field, there is ambiguity about the training background for someone with the title of "data scientist" due to lack of consistent standards. They may or may not have a graduate degree, and their degree (if they have one) could be in just about any STEM field. My understanding is that this is usually not the case for statisticians; it's not quite as robust a title as something like "medical doctor", but someone with a software engineering bachelor's typically isn't qualified to be a statistician, though they can be a data scientist. For a statistician, we can safely assume this individual has at least a master's-level understanding and formal training in statistics (theoretical and applied). On the other hand, fewer assumptions are safe for describing a data scientist. All this to say, the lack of precision for the definition of a "data scientist" makes comparisons challenging.


## How do you view yourself in relation to these two areas?

Seeing as how I'm in a statistics program, I relate more to statisticians, though I don't consider myself a statistician (yet!). While social and economic pressures may categorize me more as a data scientist (depending on where I work, especially without a doctorate I think), I place more value in learning to think and reason as a statistician. Even though I may engage in several activities that a data scientist does (e.g. programming, computing, learning business/contextual knowledge, etc), I am learning some of the mathematical theory for how the statistics I employ actually work. Analogously, someone can learn how to drive a car without knowing how a car works. This isn't a problem, per say, but when the car breaks down you won't know how to fix it, and attempts to do so could make the issue worse! Likewise, a data scientist may or may not know how the math behind their functions work, but a statistician should always have at least a basic understanding of this. I value this conceptually deeper understanding for many reasons, and I believe it ultimately tends to produce more careful, accurate, and useful results in professional and personal settings.

## Example R code

Below is an example of code examining petal width in the Iris dataset. We assign 'y' to this density, and first examine its class. 
```{r}
y <- density(iris$Petal.Width)

class(y)
```

The type of y:
```{r}
typeof(y)
```

The structure of y:
```{r}
str(y)
```

Below is a plot of y:
```{r}
#| echo: false
plot(y)
```
